How young children integrate information sources to infer the meaning of words

### 2021-02-26 paper/manuscript.Rmd
add AIS analysis files
clean repo for revision

abstract: |
  Before formal education begins, children typically acquire a vocabulary of thousands of words. This learning process requires the use of many different information sources in their social environment, including the context in which they hear words used and their current state of knowledge. How is this information integrated? We specify a developmental model according to which children consider information sources in an age-specific way and integrate them via Bayesian inference. This model accurately predicted 2-to-5 year-old children’s word learning across a range of experimental conditions. Model comparison suggests that the central locus of development is an increased sensitivity to individual information sources, rather than changes in  integration ability. This work presents a quantitative developmental theory of information integration during language learning, and illustrates how formal models can be used to make a quantitative test of the predictive and explanatory power of competing theories.

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf

@@

Human communicative abilities are unrivaled in the animal kingdom[@taylor2016language; @tomasello2008origins; @fitch2010evolution]. Language -- in whatever modality -- is the medium that allows humans to collaborate and coordinate in species-unique ways, making it the bedrock of human culture and society [@smith1997major]. Thus, to absorb the culture around them and become functioning members of society, children need to learn language [@tomasello2018natural]. A central problem in language learning is referent identification: To acquire the conventional symbolic relation between a word and an object, a child must determine the intended referent of the word. However, there is no unique cue to reference that can be used across all situations [@frank2013social]. Instead, referents can only be identified inferentially by reasoning about the speaker’s intentions [@bohn2019pervasive; @bruner1983child; @clark2009first; @tomasello2009constructing]. That is, the child has to infer what the speaker is communicating about based on information sources in the utterance’s social context.

@@

When information integration is studied directly, the focus is mostly on how children interpret an ambiguous referring expression in light of social-contextual information [@ganea2007infants; @graham2017words; @grosse201021; @jara2020social; @khu2020preschoolers; @matthews2006effect; @nilsen2009preschoolers; @kachel2021young]. In one classic study [@nadig2002evidence], children faced a 2 x 2 display with a ball, a pen and two glasses in it. The speaker, sitting on the opposite side of the display, saw only three of the four compartments: the ball, the pen, and one of the glasses. When the speaker asked for "the glass", children had to integrate the semantics of the utterance with the speaker’s perspective to correctly infer which of the glasses the speaker was referring to. This study advanced our understanding by documenting that preschoolers use both information sources, a finding confirmed by a variety of other work [@graham2017words; @khu2020preschoolers; @nilsen2009preschoolers]. Yet these studies do not specify -- or test -- the process by which children integrate different information sources. When interpreting their findings, work in this tradition refers to social-pragmatic theories of language use and learning [@grice1991studies; @sperber2001relevance; @clark1996using; @clark2009first; @tomasello2009constructing], all of which assume that information is integrated as part of a social inference process, but none of which clearly defines the process. As a consequence, we have no explicit and quantitative theory of how different information sources (and word learning mechanisms) are integrated.

We present a theory of this integration process. Following social-pragmatic theories of language learning [@clark2009first; @tomasello2009constructing], our theory is based on the following premises: information sources serve different functional roles but are combined as part of an integrated social inference process [@clark1996using; @grice1991studies; @shafto2012learning; @sperber2001relevance]. Children use all available information to make inferences about the intentions behind a speaker’s utterance, which then leads them to correctly identify referents in the world and learn conventional word–object mappings. We formalize the computational steps that underlie this inference process in a cognitive model [@frank2012predicting; @goodman2016pragmatic; @oberauer2019addressing]. In the remainder of this paper, we rigorously test this theory by asking how well it serves the two purposes of any psychological theory: prediction and explanation [@shmueli2010explain; @yarkoni2017choosing]. First, we use the model to make quantitative predictions about children’s behavior in new situations -- predictions we test against new data. This form of model testing has been successfully used with adults [@frank2012predicting; @lake2015human] and here we extend it to children. Next, we quantify how well the model explains the integration process by comparing it to alternative models that make different assumptions about *whether* information is integrated, *how* it is integrated, and how the integration process *develops.*

@@

Our *rational integration model* arbitrates between information sources via Bayesian inference (see Fig. 1f for model formulae). A listener ($L_{1}$) reasons about the referent of a speaker’s ($S_{1}$) utterance. This reasoning is contextualized by the prior probability of each referent $\rho$ -- which we take to be a function of the common ground shared between the listener and the speaker. To decide between referents, the listener ($L_{1}$) reasons about what a rational speaker ($S_{1}$) with informativeness $\alpha$ would say given an intended referent. This speaker is assumed to compute the informativity for each available utterance and then choose the most informative one. The informativity of each utterance is given by imagining which referent a listener, who interprets words according to their literal semantics (what we call a literal listener, $L_{0}$), would infer upon hearing the utterance. Naturally, this reasoning depends on what kind of semantic knowledge (for object j) $\theta$~j~ the speaker ascribes to the (literal) listener.

@@

```{r fig1, include = T, fig.align = "center", fig.cap = "Experimental task and model. (a and b) Screenshots from the experimental task. (i) The speaker encounters one object and then leaves the scene (ii). While the speaker is away, a second object appears (iii), when returning, the speaker uses a novel word to request an object (iv). Sections (i) to (iii) establish common ground between the speaker and the listener, in that one object is new in context (green). The request in (iv) licenses an inference based on expectations about how informative speakers are (red). Listeners’ semantic knowledge enters the task because the identity of the known object on one of the tables is varied from well-known objects like a duck to relatively unfamiliar objects like a chess pawn (total of 12 objects -- blue). (a) shows the version of the experiment in which common ground information is congruent (i.e., point to the same object) with speaker informativeness and (b) shows the incongruent condition. The congruent and incongruent conditions are each paired with the 12 known objects, resulting in 24 unique conditions. (c to e) show developmental trajectories for each information source, estimated from separate experiments (see supplement). (f) gives the model equation for the rational integration model and links information sources to model parameters.", out.width="100%"}
knitr::include_graphics("./figures/figure1.png")
```

To test the predictive and explanatory power of our model we designed a world-learning experiment in which we jointly manipulated the three information sources (Fig. 1). Children interacted on a tablet computer with a series of storybook speakers [@frank2016using]. This situation is depicted in Fig. 1a iv, in which a speaker (here, a frog) appears with a known object (a duck, left) and an unfamiliar object (the diamond-shaped object, right). The speakers used a novel word (e.g., "wug") in the context of two potential referents, and then the child was asked to identify a new instance of the novel word, testing their inference about the speaker’s intended referent. To vary the strength of the child’s inference, we systematically manipulated the familiarity of the known object (from e.g., the highly familiar "duck" to the relatively unfamiliar "pawn") and whether the familiar or novel object was new to the speaker (meaning that it was not part of common ground).

@@

We tested the model in its ability to predict 2 - 5 year-old children’s judgments about word meaning. We estimated children’s (N=148) developing sensitivity to individual information sources in two separate experiments (Experiments 1 and 2; see Fig. 1c-e). We then generated parameter-free *a priori* model predictions (developmental trajectories) representing the model’s expectations about how children should behave in a new situation in which all three information sources had to be integrated. We generated predictions for 24 experimental conditions: 12 objects of different familiarities (requiring different levels of semantic knowledge), with novelty either conflicting or coinciding; Fig. 1. We compared these predictions to newly collected data from N = 220 children from the same age range (Experiment 3). All procedures, sample sizes and analysis were pre-registered (see methods).

The results showed a very close alignment between model predictions and the data across the entire age range. That is, the average developmental trajectories predicted by the model resembled the trajectories found in the data (Fig. S6). For a more quantitative analysis, we binned predictions and data by child age (in years) and correlated the two. We found a high correlation, with the model explaining 79% of the variance in the data (Fig. 2a). These results support the assumption of the model that children integrate three all available information sources. However, it is still possible that simpler models might make equally good – or even better – predictions.

@@

We found little support for these heuristic models (Fig. 2b). When using Bayesian model comparison via marginal likelihood of the data [@lee2014bayesian], we find that the data was several orders of magnitude more likely under the rational integration model compared to any of the lesioned models (Fig. 2). Figure 2c exemplifies the differences between the models: all heuristic models systematically underestimate children’s performance in the congruent condition. Thus, even when the information sources are redundant (i.e. they all point to the same referent), children’s inferences are notably strengthened by each of them. In the incongruent condition, the no word knowledge model underestimates performance, because it does not differentiate between the different familiar objects, and in the case of a highly familiar word such as duck, underestimates the strength of the mutual exclusivity inference and its compensatory effect. The no speaker informativeness completely ignores this inferences, which leads to even worse predictions. On the contrary, the no common ground model overestimates performance because it ignores the dampening effect of common ground pointing to a different referent. Taken together, we conclude that children considered all available information sources.

@@

The rational integration model assumes that all information sources enter into a joint inference process, but alternative integration processes are conceivable and might be consistent with the data. For example, the "bag of tricks" [@bloom2002children] hypothesis mentioned in the introduction could be re-phrased as a modular integration process: children might compute independent inferences based on subsets of the available information and then integrate them in a post-hoc manner by weighting them according to some parameter. This view would allow for the possibility that some information sources are considered to be more important than others. In other words, children might be biased towards some information sources. We formalized this alternative view as a *biased integration model*. This model assumes that semantic knowledge and expectations about speaker informativeness enter into one inference (mutual exclusivity inference [@markman1988children; @lewis2020-me; @clark1987principle]) while common ground information enters into a second one. The outcomes of both processes are then weighted according to the parameter $\phi$. Like the rational integration model, this model takes in all available information sources in an age-sensitive way and assumes that they are integrated. The only difference lies in the nature of the integration process: the biased integration model privileges some information sources of others in an ad-hoc manner.

The parameter $\phi$ in the biased integration model is unknown ahead of time and has to be estimated based on the experimental data. That is, in contrast to the rational integration model, the biased integration model has an additional degree of freedom which does not allow us to make *a priori* predictions about new data in the way we describe above. For a fair comparison, we therefore constrained the parameters in the rational integration model by the data from Experiment 3 as well. As a consequence, both models use all the data available to constrain their parameters (see Fig. S4).

@@

The biased integration model makes reasonable predictions and explains 78% of the variance in the data (Fig. 3b). The parameter $\phi$ -- indicating the bias to one of the inferences -- was estimated to favor the mutual exclusivity inference (Maximum A-Posteriori estimate = `r phi%>%pull(mode)`; 95% highest density interval (HDI): `r phi%>%pull(lci)` - `r phi%>%pull(uci)`, see Fig. 3d). However, the rational integration model presented a much better fit to the data, both in terms of correlation and the marginal likelihood of the data (Fig. 3). Figure 3e exemplifies the difference between the models: the biased integration model puts extra weight on the mutual exclusivity inference and thus fails to capture performance when this inference is weak compared to the common ground inference -- such as in the congruent condition for younger children. Thus, a fully integrated -- as opposed to a modular and biased -- integration process explained the data better.

@@

The developmental bias model also explains a substantial portion of the variance in the data: 78% (Fig. 3c). The estimated developmental trajectory for the bias parameter $\phi$ suggests that younger children put a stronger emphasis on common ground information, while older children rely more on the mutual exclusivity inference (Fig. 3d). The relative importance of the two inferences seems to switch at around age 3. Yet again, when we directly compare the competitor models, we find that the data is several orders of magnitude more likely under the rational integration model (Fig. 3). Looking at Figure 3e, we can see that the developmental bias model tends to underestimate children’s performance because the supportive interplay between the different inferences is constrained.

Finally, when we compared the rational integration model with its parameters not constrained by the experimental data (i.e. the "pure" prediction model described above), it also outperformed both bias models (see Fig. S10 and Tab. S6). Taken together, these results show that the integration process is best described as fully integrated and stable over time.

@@

Our work extends the study of information integration during language comprehension [@tanenhaus1995integration; @kamide2003integration; @hagoort2004integration; @ozyurek2007line; @heller2016perspective] to pragmatics. Our model is based on classic social-pragmatic theories on language use and comprehension [@clark1996using; @grice1991studies; @sperber2001relevance; @tomasello2009constructing]. As a consequence, instead of assuming that different information sources feed into separate word-learning mechanisms, we assume that all of these information sources play a functional role in an integrated social inference process. Our model goes beyond previous theoretical and empirical work by specifying the computations that underlie this inference process. Furthermore, we present a substantive theory about how this integration process develops: We assume that children become increasingly sensitive to different information sources, but that the way these information sources are integrated remains the same. We used this model to predict and explain children’s information integration in a new word learning paradigm in which they had to integrate (1) their assumptions about informative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge.

We found that this rational integration model made accurate quantitative predictions across a range of experimental conditions both when information sources were aligned and were in conflict. Predictions from the model better explained the data compared to lesioned models which assumed that children ignore one of the information sources, suggesting that children used all available information. To test the explanatory power of the model -- how well it explains the process by which information is integrated -- we formalized an alternative, modular, view. According to the biased integration model, children use all available information sources but compute separate inferences based on a subset of them. Integration happens by weighing the outcomes of these separate inferences by some parameter. Finally, we tested an alternative view on the development of the integration process. According to the developmental bias model, the importance of the different information sources changed with age. In both cases, the rational integration model provided a much better fit to the data, suggesting that the integration process remains stable over time.

The rational integration model is derived from a more general framework for pragmatic inference, which has been used to explain a wide variety of phenomena in adults’ language use and comprehension [@monroe2017colors; @wang2016learning; @tessler2019language; @yoon_frank_tessler_goodman_2018;@frank2012predicting; @goodman2016pragmatic]. Thus, it can be generalized in a natural way to capture word learning in contexts that offer more, fewer, or different types of information. For example, in a recent study with adults and children, we manipulated expectations about informativeness and common ground in a different way. Using a structurally similar model [@bohn_tessler_merrick_frank_2019], we also found a close alignment between model predictions and the data. The flexibility of this modeling framework stems from its conceptualization of human communication as a form of rational social action. As such, it connects to computational and empirical work that tries to explain social reasoning by assuming that humans expect each other to behave in a way that maximizes the benefits and minimizes the cost associated with actions [@bridgers2020young; @jara2016naive; @jara2020social].

Our model and empirical paradigm provide a foundation on which to test deeper questions about language development. First, our findings should be replicated in children from different cultural backgrounds, learning different languages [@nielsen2017persistent]. In such studies, we would not expect our results to replicate in a strict sense; that is, we would not expect to see the same developmental trajectories in all cultures and languages. Substantial variation is much more likely. Nevertheless, our model provides a way to think about how to reconcile cross-cultural variation with a shared cognitive architecture: We predict differences in how sensitive children are to the individual information sources at different ages, but similarities in how information is integrated [@bohn2019pervasive]. In computational terms, we assume a universal architecture that specifies the relation between a set of varying parameters. Of course, either confirmation or disconfirmation of this prediction would be informative.

Second, it would be useful to flesh out the cognitive processes that underlie reasoning about common ground. In its current form, our model does not specify how the listener arrives at the expectation that some objects are more likely referents because they are new in common ground. We assumed that common ground changes the speaker’s expectations about what are likely referents [@hanna2003effects] and we measured these expectations to inform the parameter values in our model. But computationally, the model does not differentiate between common ground information and other reasons that might make an object contextually more salient.

Finally, our model is a model of referent identification in the moment of the utterance. At the same time, the constructs made use of by our model are shaped by factors that unfold across multiple time points and contexts:  Common ground is built over the course of a conversation, and the lexical knowledge of a child is shaped across a language developmental time-scale. Our model makes use of unidimensional representations of these high-dimensional, structured processes and examines how these representations are integrated.  Connecting our model with frameworks that focus on the temporal and cross-situational aspects of word learning would elucidate further these complex, time-dependent processes [@frank2009using; @mcmurray2012word; @yurovsky2015integrative].

@@

A more detailed description of the experiments and the models can be found in the supplementary material. The experimental procedure, sample size, and analysis for each experiment were pre-registered (https://osf.io/7rg9j/registrations). Experimental procedures, model and analysis scripts can be found in an online repository (https://github.com/manuelbohn/spin). Experiments 1 and 2 were designed to estimate children’s developing sensitivity to each information source. The results of these experiments determine the parameter values in the model (see Fig. 1 c-f). Experiment 3 was designed to test how children integrate different information sources.

@@

To analyze how the manipulations in each experiment affected children’s behavior, we used generalized linear mixed models. Since the focus of the paper is on how information sources were integrated, we discuss these models in the supplementary material and focus here on the cognitive models instead. A detailed, mathematical description of the different cognitive models along with details about estimation procedures and priors can be found in the supplementary material. All cognitive models and Bayesian data analytic models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository.

@@

>>>>

abstract: |
  Before formal education begins, children typically acquire a vocabulary of thousands of words. This learning process requires the use of many different information sources in their social environment, including the context in which they hear words used and their current state of knowledge. How is this information integrated? We specify a developmental model according to which children consider information sources in an age-specific way and integrate them via Bayesian inference. This model accurately predicted 2-to-5 year-old children’s word learning across a range of experimental conditions in which they had to integrate three information sources. Model comparison suggests that the central locus of development is an increased sensitivity to individual information sources, rather than changes in  integration ability. This work presents a quantitative developmental theory of information integration during language learning, and illustrates how formal models can be used to make a quantitative test of the predictive and explanatory power of competing theories.

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_docx

@@

Human communicative abilities are unrivaled in the animal kingdom [@taylor2016language; @tomasello2008origins; @fitch2010evolution]. Language -- in whatever modality -- is the medium that allows humans to collaborate and coordinate in species-unique ways, making it the bedrock of human culture and society [@smith1997major]. Thus, to absorb the culture around them and become functioning members of society, children need to learn language [@tomasello2018natural]. A central problem in language learning is referent identification: To acquire the conventional symbolic relation between a word and an object, a child must determine the intended referent of the word. However, there is no unique cue to reference that can be used across all situations [@frank2013social]. Instead, referents can only be identified inferentially by reasoning about the speaker’s intentions [@bohn2019pervasive; @bruner1983child; @clark2009first; @tomasello2009constructing]. That is, the child has to infer what the speaker is communicating about based on information sources in the utterance’s social context.

@@

When information integration is studied directly, the focus is mostly on how children interpret or learn words in light of social-contextual information [@hollich2000breaking; @ganea2007infants; @graham2017words; @grosse201021; @jara2020social; @khu2020preschoolers; @matthews2006effect; @nilsen2009preschoolers; @kachel2021young]. In one classic study [@nadig2002evidence], children faced a 2 x 2 display with a ball, a pen and two glasses in it. The speaker, sitting on the opposite side of the display, saw only three of the four compartments: the ball, the pen, and one of the glasses. When the speaker asked for "the glass", children had to integrate the semantics of the utterance with the speaker’s perspective to correctly infer which of the glasses the speaker was referring to. This study advanced our understanding by documenting that preschoolers use both information sources, a finding confirmed by a variety of other work [@graham2017words; @khu2020preschoolers; @nilsen2009preschoolers]. Yet these studies do not specify -- or test -- the process by which children integrate different information sources. When interpreting their findings, work in this tradition refers to social-pragmatic theories of language use and learning [@grice1991studies; @sperber2001relevance; @clark1996using; @clark2009first; @tomasello2009constructing], all of which assume that information is integrated as part of a social inference process, but none of which clearly defines the process. As a consequence, we have no explicit and quantitative theory of how different information sources (and word learning mechanisms) are integrated.

We present a theory of this integration process. Following social-pragmatic theories of language learning [@clark2009first; @tomasello2009constructing], our theory is based on the following premises: information sources serve different functional roles but are combined as part of an integrated social inference process [@clark1996using; @grice1991studies; @shafto2012learning; @sperber2001relevance]. Children use all available information to make inferences about the intentions behind a speaker’s utterance, which then leads them to correctly identify referents in the world and learn conventional word–object mappings. We formalize the computational steps that underlie this inference process in a cognitive model[@frank2012predicting; @goodman2016pragmatic; @oberauer2019addressing]. In contrast to earlier modelling work, we treat word learning as the outcome of a social inference process instead of a cross-situational[@fazly2010probabilistic; @frank2009using] or principle-based learning process [@xu2007word]. In the remainder of this paper, we rigorously test this theory by asking how well it serves the two purposes of any psychological theory: prediction and explanation [@shmueli2010explain; @yarkoni2017choosing]. First, we use the model to make quantitative predictions about children’s behavior in new situations -- predictions we test against new data. This form of model testing has been successfully used with adults [@frank2012predicting; @lake2015human] and here we extend it to children. Next, we quantify how well the model explains the integration process by comparing it to alternative models that make different assumptions about *whether* information is integrated, *how* it is integrated, and how the integration process *develops*. Alternative models either assume that children ignore some information sources or -- in line with a “bag of tricks” approach -- they assume that children compute isolated inferences and then weigh their outcome in a post-hoc manner.

@@

Our *rational integration model* arbitrates between information sources via Bayesian inference (see Fig. 1f for model formulae). A listener ($L_{1}$) reasons about the referent of a speaker’s ($S_{1}$) utterance. This reasoning is contextualized by the prior probability of each referent $\rho$. We treat $\rho$ as a conversational prior which originates from the common ground shared between the listener and the speaker. This interpretation follows from the social nature of our experiments (see below). From a modelling perspective, $\rho$ can be (and also also has been) used to capture non-social aspects of a referent, for example its visual salience[@frank2012predicting]. To decide between referents, the listener ($L_{1}$) reasons about what a rational speaker ($S_{1}$) with informativeness $\alpha$ would say given an intended referent. This speaker is assumed to compute the informativity for each available utterance and then choose the most informative one. The informativity of each utterance is given by imagining which referent a listener, who interprets words according to their literal semantics (what we call a literal listener, $L_{0}$), would infer upon hearing the utterance. Naturally, this reasoning depends on what kind of semantic knowledge (for object j) $\theta$~j~ the speaker ascribes to the (literal) listener.

@@

(ref:figlab1) **Experimental task and model**. (a and b) Screenshots from the experimental task. (i) The speaker encounters one object and then leaves the scene. (ii) While the speaker is away, (iii) a second object appears, (iv) when returning, the speaker uses a novel word to request an object. Sections (i) to (iii) establish common ground between the speaker and the listener, in that one object is new in context (red). The request in (iv) licenses an inference based on expectations about how informative speakers are (gold). Listeners’ semantic knowledge enters the task because the identity of the known object on one of the tables is varied from well-known objects like a duck to relatively unfamiliar objects like a chess pawn (total of 12 objects -- blue). (a) shows the condition of the experiment in which common ground information is congruent (i.e., point to the same object) with speaker informativeness and (b) shows the incongruent condition. The congruent and incongruent conditions are each paired with the 12 known objects, resulting in 24 unique conditions. Developmental trajectories are shown for (c) sensitivity to common ground, (d) speaker informativeness and (e) semantic knowledge, estimated based on separate experiments (see main text). (f) gives the model equation for the rational integration model and links information sources to model parameters.

```{r fig1, include = T, fig.align = "center", fig.cap = "(ref:figlab1)", out.width="100%"}
knitr::include_graphics("./figures/figure1.png")
```

To test the predictive and explanatory power of our model we designed a word-learning experiment in which we jointly manipulated the three information sources (Fig. 1). Children interacted on a tablet computer with a series of storybook speakers [@frank2016using]. This situation is depicted in Fig. 1a iv, in which a speaker (here, a frog) appears with a known object (a duck, left) and an unfamiliar object (the diamond-shaped object, right). The speakers used a novel word (e.g., "wug") in the context of two potential referents, and then the child was asked to identify a new instance of the novel word, testing their inference about the speaker’s intended referent. To vary the strength of the child’s inference, we systematically manipulated the familiarity of the known object (from e.g., the highly familiar "duck" to the relatively unfamiliar "pawn") and whether the familiar or novel object was new to the speaker (meaning that it was not part of common ground).

@@

We tested the model in its ability to predict 2 - 5 year-old children’s judgments about word meaning. We estimated children’s (N=148) developing sensitivity to individual information sources in two separate experiments (Experiments 1 and 2; see Fig. 1c-e). In Experiment 1, we jointly estimated children’s sensitivity to informativeness and their semantic knowledge. In Experiment 2, we estimated sensitivity to common ground. We then generated parameter-free *a priori* model predictions (developmental trajectories) representing the model’s expectations about how children should behave in a new situation in which all three information sources had to be integrated. We generated predictions for 24 experimental conditions: 12 objects of different familiarities (requiring different levels of semantic knowledge), with novelty either conflicting or coinciding; Fig. 1. We compared these predictions to newly collected data from N = 220 children from the same age range (Experiment 3). All procedures, sample sizes and analysis were pre-registered (see methods).

The results showed a very close alignment between model predictions and the data across the entire age range. That is, the average developmental trajectories predicted by the model resembled the trajectories found in the data (Fig. S6). For a more quantitative analysis, we binned predictions and data by child age (in years) and correlated the two. We found a high correlation, with the model explaining 79% of the variance in the data (Fig. 2a). These results support the assumption of the model that children integrate three all available information sources. However, it is still possible that simpler models might make equally good – or even better – predictions. For example, work on children’s use of statistical information during word learning showed that their behaviour was best explained by a model which selectively ignored parts of the input [@gagliardi2017modeling].

@@

We found little support for these heuristic models (Fig. 2b). When using Bayesian model comparison via marginal likelihood of the data [@lee2014bayesian], we find that the data was several orders of magnitude more likely under the rational integration model compared to any of the lesioned models (Fig. 2). Figure 2c exemplifies the differences between the models: all heuristic models systematically underestimate children’s performance in the congruent condition. Thus, even when the information sources are redundant (i.e. they all point to the same referent), children’s inferences are notably strengthened by each of them. In the incongruent condition, the no word knowledge model underestimates performance, because it does not differentiate between the different familiar objects, and in the case of a highly familiar word such as duck, underestimates the strength of the mutual exclusivity inference and its compensatory effect. The no speaker informativeness completely ignores this inferences, which leads to even worse predictions. On the contrary, the no common ground model overestimates performance because it ignores the dampening effect of common ground favoring a different referent. Taken together, we conclude that children considered all available information sources.

@@

The rational integration model assumes that all information sources enter into a joint inference process, but alternative integration processes are conceivable and might be consistent with the data. For example, the "bag of tricks" [@bloom2002children] hypothesis mentioned in the introduction could be re-phrased as a modular integration process: children might compute independent inferences based on subsets of the available information and then integrate them in a post-hoc manner by weighting them according to some parameter. This view would allow for the possibility that some information sources are considered to be more important than others. In other words, children might be biased towards some information sources. We formalized this alternative view as a *biased integration model*. This model assumes that semantic knowledge and expectations about speaker informativeness enter into one inference (mutual exclusivity inference [@markman1988children; @lewis2020-me; @clark1987principle]) while common ground information enters into a second one. The outcomes of both processes are then weighted according to the parameter $\phi$. Like the rational integration model, this model takes in all available information sources in an age-sensitive way and assumes that they are integrated. The only difference lies in the nature of the integration process: the biased integration model privileges some information sources over others in an ad-hoc manner.

The parameter $\phi$ in the biased integration model is unknown ahead of time and has to be estimated based on the experimental data. That is, through Experiment 1 and 2 alone, we do not learn anything about the relative importance of the information sources. As a consequence -- and in contrast to the rational integration model -- the biased integration model does not allow us to make *a priori* predictions about new data in the way we describe above. For a fair comparison, we therefore constrained the parameters in the rational integration model by the data from Experiment 3 as well. As a consequence, both models estimate their parameters using all the data available in a fully Bayesian manner (see Fig. S4).

@@

The biased integration model makes reasonable predictions and explains 78% of the variance in the data (Fig. 3b). The parameter $\phi$ -- indicating the bias to one of the inferences -- was estimated to favor the mutual exclusivity inference (Maximum A-Posteriori estimate = `r phi%>%pull(mode)`; 95% highest density interval (HDI): `r phi%>%pull(lci)` - `r phi%>%pull(uci)`, see Fig. 3d). However, the rational integration model presented a much better fit to the data, both in terms of correlation and the marginal likelihood of the data (Fig. 3). When constrained by the data from all experiments, the rational integration model explains 87% of the variance in the data. Fig. 3e exemplifies the difference between the models: the biased integration model puts extra weight on the mutual exclusivity inference and thus fails to capture performance when this inference is weak compared to the common ground inference -- such as in the congruent condition for younger children. As a result, a fully integrated -- as opposed to a modular and biased -- integration process explained the data better.

@@

The developmental bias model also explains a substantial portion of the variance in the data: 78% (Fig. 3c). The estimated developmental trajectory for the bias parameter $\phi$ suggests that younger children put a stronger emphasis on common ground information, while older children rely more on the mutual exclusivity inference (Fig. 3d). The relative importance of the two inferences seems to switch at around age 3. Yet again, when we directly compare the competitor models, we find that the data is several orders of magnitude more likely under the rational integration model (Fig. 3). Looking at Figure 3e, we can see that the developmental bias model tends to underestimate children’s performance because the supportive interplay between the different inferences is constrained. In the biased models, the overall inference can only be as strong as the strongest of the components -- in the rational integration model, the components interact with one another, enabling a stronger overall inference.

@@

Previous work on  the study of information integration during language comprehension focused on how adults combine perceptual, semantic or syntactic information [@tanenhaus1995integration; @kamide2003integration; @hagoort2004integration; @ozyurek2007line; @heller2016perspective]. Our work extends this work to the development of pragmatics. Our model is based on classic social-pragmatic theories on language use and comprehension [@clark1996using; @grice1991studies; @sperber2001relevance; @tomasello2009constructing]. As a consequence, instead of assuming that different information sources feed into separate word-learning mechanisms (the “bag of tricks” view), we assume that all of these information sources play a functional role in an integrated social inference process. Our model goes beyond previous theoretical and empirical work by specifying the computations that underlie this inference process. Furthermore, we present a substantive theory about how this integration process develops: We assume that children become increasingly sensitive to different information sources, but that the way these information sources are integrated remains the same. We used this model to predict and explain children’s information integration in a new word learning paradigm in which they had to integrate (1) their assumptions about informative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge.

We found that this rational integration model made accurate quantitative predictions across a range of experimental conditions both when information sources were aligned and were in conflict. Predictions from the model better explained the data compared to lesioned models which assumed that children ignore one of the information sources, suggesting that children used all available information. To test the explanatory power of the model -- how well it explains the process by which information is integrated -- we formalized an alternative, modular, view. According to the biased integration model, children use all available information sources but compute separate inferences based on a subset of them. Integration happens by weighing the outcomes of these separate inferences by some parameter. Finally, we tested an alternative view on the development of the integration process. According to the developmental bias model, the importance of the different information sources changed with age. In both cases, the rational integration model provided a much better fit to the data, suggesting that the integration process remains stable over time. That is, there is developmental continuity and therefore no qualitative difference in how a 2-year-old integrates information compared to a 4-year-old.

The rational integration model is derived from a more general framework for pragmatic inference, which has been used to explain a wide variety of phenomena in adults’ language use and comprehension [@monroe2017colors; @wang2016learning; @tessler2019language; @yoon2020polite; @frank2012predicting; @goodman2016pragmatic; @savinelli2017modeling]. Thus, it can be generalized in a natural way to capture word learning in contexts that offer more, fewer, or different types of information. For example, non-verbal aspects of the utterance (e.g. eye-gaze or gestures) can affect children’s mutual exclusivity inference [@paulus2014conflicting; @graham2010role; @gangopadhyay2020role; @jaswal2006learning; @grassmann2010young]. As a first step in this direction, we recently studied how adults and children integrate non-verbal utterances with common ground[@bohn_tessler_merrick_frank_2019]. Using a structurally similar model, we also found a close alignment between model predictions and the data. The flexibility of this modeling framework stems from its conceptualization of human communication as a form of rational social action. As such, it connects to computational and empirical work that tries to explain social reasoning by assuming that humans expect each other to behave in a way that maximizes the benefits and minimizes the cost associated with actions [@bridgers2020young; @jara2016naive; @jara2020social].

Our model and empirical paradigm provide a foundation on which to test deeper questions about language development. First, our findings should be replicated in children from different cultural backgrounds, learning different languages [@nielsen2017persistent]. In such studies, we would not expect our results to replicate in a strict sense; that is, we would not expect to see the same developmental trajectories in all cultures and languages. Substantial variation is much more likely. Studies on children’s pragmatic inferences in different cultures have documented similar[@su2015interpretation; @zhao2019development] and different[@fortier_kellier_flecha_frank_2018] developmental trajectories. Nevertheless, our model provides a way to think about how to reconcile cross-cultural variation with a shared cognitive architecture: We predict differences in how sensitive children are to the individual information sources at different ages, but similarities in how information is integrated [@bohn2019pervasive]. In computational terms, we assume a universal architecture that specifies the relation between a set of varying parameters. Of course, either confirmation or disconfirmation of this prediction would be informative.

Second, it would be useful to flesh out the cognitive processes that underlie reasoning about common ground. The basic assumption that common ground changes interlocutors’ expectations about what are likely referents[@hanna2003effects] has been used in earlier modelling work on the role of common ground in reference resolution [@heller2016perspective]. Here we went one step further and measured the strength of these expectations to inform the parameter values in our model. However, in its current form, our model treats common ground as a conversational prior and does not specify how the listener arrives at the expectation that some objects are more likely referents because they are new in common ground. That is, computationally, our model does not differentiate between common ground information and other reasons that might make an object contextually more salient. An interesting starting point to overcome this shortcoming would be modelling work on the role of common ground in conversational turn taking [@anderson2021tell].

Finally, our model is a model of referent identification in the moment of the utterance. At the same time, the constructs made use of by our model are shaped by factors that unfold across multiple time points and contexts: Common ground is built over the course of a conversation, and the lexical knowledge of a child is shaped across a language developmental time-scale. Even speaker informativeness could be imagined to vary over time following repeated interactions with a particular speaker. What is more, assessing speaker informativeness is unlikely to be the outcome of a single, easy-to-define process. The expectations about informative communication that we take it to represent are probably the result of the interplay between multiple social and non-social inference processes. Thus, our model makes use of unidimensional representations of these high-dimensional, structured processes and examines how these representations are integrated. Connecting our model with other frameworks that focus on the cognitive, temporal and cross-situational aspects of word learning would elucidate further these complex processes [@frank2009using; @mcmurray2012word; @yurovsky2015integrative].

@@

A more detailed description of the experiments and the models can be found in the supplementary material. The experimental procedure, sample sizes, and analysis for each experiment were pre-registered (https://osf.io/7rg9j/registrations). Experimental procedures, model and analysis scripts can be found in an online repository (https://github.com/manuelbohn/spin). Experiments 1 and 2 were designed to estimate children’s developing sensitivity to each information source. The results of these experiments determine the parameter values in the model (see Fig. 1 c-f). Experiment 3 was designed to test how children integrate different information sources.

@@

To analyze how the manipulations in each experiment affected children’s behavior, we used generalized linear mixed models. Since the focus of the paper is on how information sources were integrated, we discuss these models in the supplementary material and focus here on the cognitive models instead. A detailed, mathematical description of the different cognitive models along with details about estimation procedures and priors can be found in the supplementary material. All cognitive models and Bayesian data analytic models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository. Information about priors for parameter estimation and Markov chain Monte Carlo settings can also be found in the supplementary information and the online repository.

@@

# Acknowledgements

M. Bohn received funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Sklodowska-Curie grant agreement no. 749229. M. H. Tessler was funded by the National Science Foundation SBE Postdoctoral Research Fellowship Grant No. 1911790. M. C. Frank was supported by a Jacobs Foundation Advanced Research Fellowship and the Zhou Fund for Language and Cognition.

# Author Contributions

M. Bohn, M.H. Tessler and M.C. Frank conceptualized the study, M. Merrick collected the data, M. Bohn and M.H. Tessler analyzed the data, M. Bohn, M. H. Tessler and M.C. Frank wrote the manuscript, all authors approved the final version of the manuscript.
